Visualizing Multivariate PDFs

### 2016-08-25 sections/seperable.tex
figs compiles using standalone

@@

\begin{figure}
\includegraphics{figs/boxplot.png}
\caption{Tukey's 1977 example of the box plot (exhibit 6 of chapter 2 from
Exploratory Data Analytics\cite{tukey1977}) illustrates how the box plot shows
the distribution of topology of states and
the distribution of volcanos. The strength of the box and whisker is that the
reader can quickly compare the two distributions and learn, for example, that
volcanos tend to span a shorter range of average heights, but that the extremes
are much further from the center. Essentially, the distributions act as
references for the other displayed distributions.}
\label{fig:boxplot}
\end{figure}

In Tukey's introduction to the boxplot \cite{tukey1977}, shown in
figure~\ref{fig:boxplot}, the box and whisker plot takes the distributional information of a histogram and encodes the state heights between the
25th and 75th percentiles as a box, the median height as a line cutting across
the heights, and the extreme state heights as lines coming out of the
boxes. While a histogram would also show that the distribution of state heights has a fat right tail,
the power of the box and whisker plot is in encoding this information such that
it's trivial to compare to it's neighboring distribution of volcano
heights because they can share the Y-axis. As the number of distributions grow,
this technique becomes invaluable because the number of histograms that can be
placed next to each other or overlayed as pdfs can quickly become
unweildy.  Many others authors have kept the basic structure of the boxplot,
but exploited the structure to convey more information about the data. Authors used
different quantile levels \cite{Hyndman}s or measures or extremes\cite{Frigge,
carter}, proposed assymetric whiskers \cite{Rousseuw}, and otherwise incorporated skewness, kurtosis, and other descriptive distributional statsics \cite{ Aslam, choon, Marmelejo}.

\begin{figure}
%%\includegraphic{Mcgill} #annottate figure with highlighted line for the notches!
\label{fig:notched}
\end{figure}
The box-plot was then expanded out from descriptive statistics to facilitatting
inferential deduction through the notched boxplot. \cite{mcgill1978} The notched boxplot
arose as an expansion of indicating sample size through the width of the box,
which is a technique that only works well for relatively small sample sizes ($$\<1000$$), at larger scales the widths are about equal. Instead, notches, as seen
in Figure~\ref{fig:notched}, are used to indicate the confidence level of the
median for that box. The medians are statistically different from each other at
the 95\% confidence level if the notches do not overlap. In
Figure~\ref{fig:notched}, all the notches overlap %%annotate figure with what
%%overlapping region means
and so the difference in telephone bills is not-significant, but there's a
statistically significant difference between the 11-15 and over 15 groups. Notches may extend
beyond the box edges, and sometimes the width technique is incorporated back
into the notched boxpplot such that the confidence can be grounded in sample
size. The notches in Figure~\ref{fig:notched} are computed assuming a gaussian
distribution, but many authors \cite{x,y,z} discuss ways to compute notches for
non-gaussian distributions.

@@

@@

While boxplots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a timeseries
curve (where the connections are as important as the individual points) or
spatial region. These visualizations rely on the idea of ordering the
functions such that there's some measure of how close these functions are to
each other-functional depth\cite{fabrero2007}, bivariate score depth\cite{hyndman2009}, and
bivariate kernel density estimation\cite{scott92} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. For small values, the ordered curves can then be colored based on the
ordering metric \cite{hyndman2009}, but this method is unweildy for larger
ensembles.

@@

\begin{figure}
%%\includegraphics{figs/countour_weather.png}
\label{fig:countour}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the uncertainity via a band depth ordering metric. Each ensemble members band depth is computed as sum of the probabilities that the observations in any given ensemble fall within the max-min envelope defined by any two other ensembles. The bands are then sorted by band depth such that the median is the ensemble with about 50\% of its members withen all envelopes formed by other bands (so most centered). Outer bands are chosen according to the task at hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature visualizations. %
The authors argue that countour boxplots are an improved visual idiom over the traditional spagehttie plots seen in \ref{fig:countour} because as seen in (fig), the spagehetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in (fig3), the authors remove most of the bands and instead visualize an outliser envelope(light gray) and a more central envelope(dark gray). It retains much of the information of the spagethhit, but removes the visual noise of the lines.

>>>>

\documentclass[../main.tex]{subfiles}

\begin{document}
\section{Seperable Dimensions}
\label{sec:sepdim}

@@

\subsection{Discrete observations}

\begin{figure}
\includegraphics{boxplot}
\caption{Tukey's 1977 example of the box plot (exhibit 6 of chapter 2 from
Exploratory Data Analytics\cite{tukey1977}) illustrates how the box plot shows
the distribution of topology of states and
the distribution of volcanos. The strength of the box and whisker is that the
reader can quickly compare the two distributions and learn, for example, that
volcanos tend to span a shorter range of average heights, but that the extremes
are much further from the center. Essentially, the distributions act as
references for the other displayed distributions.}
\label{fig:boxplot}
\end{figure}

In Tukey's introduction to the boxplot \cite{tukey1977}, shown in
Figure~\ref{fig:boxplot}, the box and whisker plot takes the distributional information of a histogram and encodes the state heights between the
25th and 75th percentiles as a box, the median height as a line cutting across
the heights, and the extreme state heights as lines coming out of the
boxes. While a histogram would also show that the distribution of state heights has a fat right tail,
the power of the box and whisker plot is in encoding this information such that
it's trivial to compare to it's neighboring distribution of volcano
heights because they can share the Y-axis. As the number of distributions grow,
this technique becomes invaluable because the number of histograms that can be
placed next to each other or overlayed as pdfs can quickly become
unweildy.  Many others authors have kept the basic structure of the boxplot,
but exploited the structure to convey more information about the data. Authors used
different quantile levels \cite{Hyndman}s or measures or extremes\cite{Frigge,
carter}, proposed assymetric whiskers \cite{Rousseuw}, and otherwise incorporated skewness, kurtosis, and other descriptive distributional statsics \cite{ Aslam, choon, Marmelejo}.

\subsubsection{Incoporating Density}

\begin{figure}
%%\includegraphic{Mcgill} #annottate figure with highlighted line for the notches!
\label{fig:notched}
\end{figure}
The box-plot was then expanded out from descriptive statistics to facilitating
inferential deduction through the notched boxplot. \cite{mcgill1978} The notched boxplot
arose as an expansion of indicating sample size through the width of the box,
which is a technique that only works well for relatively small sample sizes ($1\<1000$), at larger scales the widths are about equal. Instead, notches, as seen
in Figure~\ref{fig:notched}, are used to indicate the confidence level of the
median for that box. The medians are statistically different from each other at
the 95\% confidence level if the notches do not overlap. In
Figure~\ref{fig:notched}, all the notches overlap %%annotate figure with what
%%overlapping region means
and so the difference in telephone bills is not-significant, but there's a
statistically significant difference between the 11-15 and over 15 groups. Notches may extend
beyond the box edges, and sometimes the width technique is incorporated back
into the notched boxpplot such that the confidence can be grounded in sample
size. The notches in Figure~\ref{fig:notched} are computed assuming a gaussian
distribution, but many authors \cite{x,y,z} discuss ways to compute notches for
non-gaussian distributions.

@@

\subsubsection{2D Distributions}

@@

\subsection{Functional `observations`}

While boxplots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a timeseries
curve (where the connections are as important as the individual points) or
spatial region. Understanding the distributions of functional observations
often relies first ordering the functions such that there's some measure of how
close these functions are to each other-functional depth\cite{fabrero2007}, bivariate score depth\cite{hyndman2009}, and
bivariate kernel density estimation\cite{scott92} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. For small values, the ordered curves can then be colored based on the
ordering metric \cite{hyndman2009}, but this method is unweildy for larger
ensembles.

@@

\begin{figure}
\includegraphics{contour_weather}
\label{fig:countour}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the uncertainity via a band depth ordering metric. Each ensemble members band depth is computed as sum of the probabilities that the observations in any given ensemble fall within the max-min envelope defined by any two other ensembles. The bands are then sorted by band depth such that the median is the ensemble with about 50\% of its members withen all envelopes formed by other bands (so most centered). Outer bands are chosen according to the task at hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature visualizations. %
The authors argue that countour boxplots are an improved visual idiom over the traditional spagehttie plots seen in \ref{fig:countour} because as seen in (fig), the spagehetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in (fig3), the authors remove most of the bands and instead visualize an outliser envelope(light gray) and a more central envelope(dark gray). It retains much of the information of the spagethhit, but removes the visual noise of the lines.  
\end{document}
